---
title: "From methods development to the clinic"
subtitle: "How to upskill your use of AI in SMARTbiomed research"
author: "Michael Sachs"
date: today
institute: "[smartbiomed.dk](https://smartbiomed.dk)"
format:
  revealjs:
    logo: SMARTbiomed_logo_left_pos.png
    theme: [default, custom.scss]
    slide-number: c/t
    progress: true
    preview-links: auto
    footer: "SMARTbiomed Annual Review"
    transition: fade
    highlight-style: github
    code-line-numbers: true
    execute:
      echo: true
      eval: false
    fig-align: center
    width: 1280
    height: 720
    scrollable: true
editor_options: 
  chunk_output_type: console
---

## Today's roadmap {.smaller}

**Four parts — ~40 minutes of content, then 15 minutes discussion**

::: {.columns}
::: {.column width="55%"}
1. How LLMs work *(theory for statisticians)*
2. The methods-to-clinic pipeline
3. Developing methods with AI: git, tests, and a simulation deep dive
4. Validate · Communicate · Govern
:::
::: {.column width="45%"}
**Opening poll** — how do you use AI today?

- Not yet
- Writing (emails, methods text, …)
- Coding (autocomplete, chat)
- Methods & analysis
- Other

:::
:::

::: notes
~2 min. Use Slido or Mentimeter for the poll; a show of hands works fine.
Set the tone: this is a working seminar, not a vendor pitch. You will hear honest failure modes.
:::

# The pipeline {background-color="#002546"}

## The pipeline view

:::: {.columns}
::: {.column width="72%"}

```{=html}
<div style="display:flex; align-items:center; gap:0.5em; font-size:0.85em; margin-top:1em;">
  <div class="stage-box">Discover<br><small>orient · literature · methods</small></div>
  <div class="stage-arrow">→</div>
  <div class="stage-box">Develop<br><small>code · simulate · test</small></div>
  <div class="stage-arrow">→</div>
  <div class="stage-box">Validate<br><small>check · diagnose · explain</small></div>
  <div class="stage-arrow">→</div>
  <div class="stage-box">Translate<br><small>write · report · tailor</small></div>
</div>
<div style="margin-top:1.2em; background:#87D1F4; color:#002546; padding:0.5em 1em; border-radius:4px; font-size:0.82em; font-weight:600; text-align:center;">
  <strong>Govern</strong> · privacy · reproducibility · verification · disclosure
</div>
```

:::
::: {.column width="28%"}
::: {.callout-tip icon=false}
**Ground rule**

AI accelerates the work.
**You** are responsible for it.
:::
:::
::::

::: notes
~2 min. Walk the four stages briefly — we will revisit each in turn.
Govern is not a stage you reach; it runs across everything.
Come back to the ground rule throughout the talk.
:::

# How LLMs work {background-color="#002546"}

## What is an LLM? A statistician's view

::: {.columns}
::: {.column width="55%"}
**A conditional probability model over text**

$$P(\text{next token} \mid \text{text so far})$$

- Text is chopped into **tokens** (~¾ word each)
- A **context window** is the max tokens it sees at once
- **Temperature** $\tau$ controls sampling sharpness:
  $\tau \to 0$ → greedy; $\tau \to \infty$ → uniform
:::
::: {.column width="45%"}
**How it's trained**

1. **Next-token prediction** on massive text corpora
2. **Instruction tuning** — follow prompts, be helpful
3. **RLHF / RLAIF** — human (or AI) preference ranking

**Why code is a strength**

Code is abundant, structured, and *correct by execution* —
a uniquely checkable signal in training data.
:::
:::

::: notes
~3 min. This is the statistician's frame: a sequence model, not a knowledge base.
Tokens and context window matter practically — large files may be truncated.
Temperature matters when you want reproducibility (set it to 0 or use a seed).
RLHF = Reinforcement Learning from Human Feedback.
:::


## Exercise - Build a lLM in 5 minutes

- Find a corpus to train the model and tokenize: 

```{r}
#| eval: true
## Works of Edgar Allan Poe, Volume I
corpus <- readLines("https://www.gutenberg.org/cache/epub/2147/pg2147.txt")
tokens <- unlist(tokenizers::tokenize_words(corpus))
```

- Train a Markov model

```{r}
#| eval: true
bigram  <- lapply(
    split(tokens[-1], tokens[-length(tokens)]),
    table
  )
unigram <- table(tokens)

model <- list(bigram = bigram, unigram = unigram)

```

## Continued

- Sample from the model

```{r}
#| eval: true
sample_next <- function(model, word) {
  counts <- model$bigram[[tolower(word)]]
  if (is.null(counts)) counts <- model$unigram
  sample(names(counts), 1, prob = counts)
}

generate <- function(model, start, n = 20) {
  out <- start
  for (i in seq_len(n)) {
    nxt <- sample_next(model, tail(out, 1))
    if (is.na(nxt)) break
    out <- c(out, nxt)
  }
  paste(out, collapse = " ")
}

generate(model, "bird", 10)
```

### Discussion

- What LLM concepts does this illustrate?
- How can this model be improved? 
- What else can it be used for? 


## Why they're confident — and sometimes wrong

::: {.columns}
::: {.column width="58%"}
**Hallucination**

> The model generates *plausible text*, not *verified truth*.

- Fabricated citations that look real *(always verify)*
- Confident wrong code that compiles but is subtly incorrect
- Invented API arguments or package functions

**Reasoning models** (o3, R1, extended thinking) trade
compute for better multi-step work, but the same failure
modes apply at higher stakes.
:::
::: {.column width="42%"}
::: {.callout-important icon=false}
**Always verify**

Every citation · every number
every derivation · every API call

Treat model output as a
**hypothesis, not a result.**
:::

::: {.callout-warning icon=false}
**Quiet deskilling**

Outsourcing judgement builds
dependency, not expertise.
Stay in the loop intentionally.
:::
:::
:::

::: notes
~2 min. Hallucination is the canonical failure mode — give a concrete example if you have one.
The key message: the model cannot be wrong and *know* it is wrong. It has no uncertainty signal it will share with you reliably.
:::

## The context window and memory {.smaller}

::: {.columns}
::: {.column width="50%"}
**What the model "sees"**

| | Typical range |
|---|---|
| Small / fast models | 8k–32k tokens |
| Standard assistants | 128k–200k tokens |
| Long-context leaders | 1M+ tokens |

A context window is **not memory** —
each new conversation starts fresh.

**What counts against it**

- Your prompt
- The entire conversation so far
- Any files or code you paste in
- Tool call results
:::
::: {.column width="50%"}
**How to work within limits**

- **Chunking**: split long documents; summarise sections before sending
- **System prompts**: keep standing instructions short and reuse across sessions
- **External memory**: project `.md` files persist context across sessions
- **Agent memory**: tools like Claude Code maintain a `CLAUDE.md` project file that persists

::: {.callout-tip icon=false}
**Practical rule**

If the conversation is getting long,
start fresh — older context degrades
response quality before it truncates.
:::
:::
:::

::: notes
~2 min. Statisticians often paste in large datasets — remind them this is expensive and may hit limits.
The "context is not memory" point surprises people. The model has no recall between sessions unless you build it in.
Agent tools (Claude Code, Codex) handle this better with persistent project files.
:::

# Stage 1: Discover & frame {background-color="#002546"}

## Discover & frame

**Orient fast in a new subfield · surface candidate methods · anchor literature work**

::: {.columns}
::: {.column width="55%"}
**What AI does well here**

- Rapid orientation: *"explain doubly-robust estimation to a statistician"*
- Surfacing related methods and their trade-offs
- Summarising abstracts, structuring a literature map
- Drafting search strings for PubMed / Web of Science

:::
::: {.column width="45%"}
::: {.callout-important icon=false}
**Big caution: confabulated citations**

The model will generate plausible
author names, journal names,
and years — that do not exist.

**Rule:** verify *every* reference
against the real source before
citing or acting on it.
:::

**Tools**

- Chat interfaces for exploration
- Elicit · Semantic Scholar for literature search
:::
:::

::: notes
~2 min. This is the safest AI use-case but also the trickiest — easy to over-trust.
Citation fabrication is not fixed in any current model. Show an example if you have one.
Specialised tools like Elicit and Semantic Scholar are safer for literature because they retrieve real documents.
:::

# Stage 2: Develop {background-color="#002546"}

## Three tiers of tools {.smaller}

Move up the tiers as the task grows in scope and complexity.

::: {.columns}
::: {.column width="33%"}
<span class="tier">Tier 1</span> **Chat interfaces**

ChatGPT · Claude · Gemini · Copilot Chat

- Paste code, ask questions
- One-shot generation
- No access to your full repo
- Good for: explaining, drafting snippets, quick checks

:::
::: {.column width="33%"}
<span class="tier">Tier 2</span> **In-IDE assistants**

GitHub Copilot · RStudio Copilot
`ellmer` · `chattr` · `gander` · `chatlas`

- Autocomplete in your editor
- **`gander` knows your data frame's columns** — context-aware suggestions
- Good for: writing functions, transforming data, iterative coding

:::
::: {.column width="33%"}
<span class="tier">Tier 3</span> **Coding agents**

Claude Code · Codex

- Repo-aware — reads your whole project
- Can run code, read errors, iterate
- Work toward a goal you set
- Good for: multi-file tasks, test-driven development, simulation scaffolding
:::
:::

::: {.callout-tip icon=false}
**Suggested on-ramp:** start with copilot in RStudio/vscode. It is the lowest-friction step up from copy-pasting to the chat window.
:::

::: notes
~2 min. Walk the three tiers as a ladder. The audience is R-heavy so emphasise the Posit ecosystem tools.
gander is a standout because it has access to the local R session — it knows the column names in your data frame.
ellmer is the foundation package the other R tools build on.
:::

## Coding agents: from autocomplete to delegate {.smaller}

::: {.columns}
::: {.column width="55%"}
**What makes agents different**

- **Repo-aware**: they read your `DESCRIPTION`, existing tests, README
- **Tool-use**: they can run R/Python, call the shell, read error messages
- **Iterative**: they retry when tests fail — the loop is automated
- **Transparent**: all changes come as a **diff** you review and commit

**Human-in-the-loop is not optional**

- Read every diff before accepting
- Run your own checks — don't trust "tests pass" without verifying
- Commit deliberately: each commit is a checkpoint you can revert to

:::
::: {.column width="45%"}
**Workflow**

```
You: "Write a function coverage() that
      computes CI coverage, with tests"

Agent: reads repo → writes function
       → writes tests → runs tests
       → iterates on failures
       → proposes diff

You:  review diff → run tests yourself
      → commit if happy
```

::: {.callout-warning icon=false}
**The agent proposes; you decide.**
Never let an agent push to `main`
without a human reviewing the diff.
:::
:::
:::

::: notes
~2 min. Key message: agents are not autopilot. They are a very capable junior collaborator who proposes all changes as reviewable diffs.
The diff-and-commit workflow is what makes agents *safe* — nothing is permanent until you commit.
:::

# Stage 2: Foundations {background-color="#002546"}

## What coding agents stand on {.smaller}

**Agents are only as reliable as the foundations they work in. These are the skills that make agent output reviewable and reversible.**

::: {.columns}
::: {.column width="50%"}
| Foundation | Why it matters for agents |
|---|---|
| **git** | every change is a diff you can inspect and revert |
| **Unit tests** | the objective signal the agent iterates against |
| **Terminal / IDE** | you run and inspect the output yourself |
| **Reproducible envs** | `renv`, Docker, seeds — results don't change between machines |
| **CI** | automated checks catch regressions across commits |
:::
::: {.column width="50%"}
::: {.callout-tip icon=false}
**Practical priority**

If you only do two things:

1. Put your project in a **git repo**
2. Write at least a few **unit tests**

Everything else helps, but these
two give you a safety net for
everything the agent touches.
:::
:::
:::

::: notes
~2 min. This slide and the next two (text skills, git, unit tests) can be folded or skimmed for audiences already familiar with the developer toolkit — see pacing notes.
:::

## Text-based computer skills {.smaller}

**Agents live in the terminal and work with plain text. A little fluency here pays outsized dividends.**

::: {.columns}
::: {.column width="50%"}
**Key skills**

- Navigate a project directory from the command line
- Read and write **plain text files** (`.md`, `.R`, `.py`, `.txt`)
- Understand what a **project root** looks like:
  ```
  my-project/
  ├── R/              # functions
  ├── analysis/       # scripts
  ├── tests/
  ├── README.md       # what is this project?
  ├── CLAUDE.md       # instructions for the agent
  └── renv.lock       # reproducible packages
  ```
:::
::: {.column width="50%"}
**Markdown files that guide agents**

| File | Purpose |
|---|---|
| `README.md` | human-facing project overview |
| `CLAUDE.md` / `agents.md` | standing instructions for the AI agent |
| `skills.md` | reusable task templates |

::: {.callout-tip icon=false}
A well-written `CLAUDE.md` is worth
hours of re-prompting — the agent
reads it at the start of every session.
:::
:::
:::

::: notes
~1.5 min. Many researchers are not comfortable in the terminal. Acknowledge this — the tools are getting better at abstracting it (Positron, RStudio terminal panel), but basic familiarity matters.
CLAUDE.md is a practical win: one file that sets context, coding conventions, and constraints for every agent session.
:::

## git: the safety net that makes agents safe {.smaller}

::: {.columns}
::: {.column width="52%"}
**The five operations you need**

| Command | What it does |
|---|---|
| `git commit` | checkpoint — a snapshot you can return to |
| `git branch` | parallel work — experiment without risk |
| `git diff` | **what changed** — what you review before committing |
| `git revert` | undo a commit cleanly |
| pull request | propose changes for review on GitHub/GitLab |

:::
::: {.column width="48%"}
**Why git makes agents safe**

> *An agent proposes work as diffs
> you review and commit —
> always visible, always reversible.*

**Workflow with an agent:**

```
agent writes → you diff → you test
→ you commit → repeat
```

If anything looks wrong:
`git checkout -- .` or `git revert`
puts you back exactly where you were.
:::
:::

::: notes
~2 min. Many statisticians have never used git. Pitch it as "undo for your whole project", not as a collaboration tool.
The key mental model: commit often — every commit is a save point. Working with an agent without git means you can't cleanly undo anything it does.
:::

## Unit tests: making "correct" checkable {.smaller}

**A test is a function that runs your code on a known input and asserts the expected output. It turns "I think this is right" into a pass/fail signal.**

::: {.columns}
::: {.column width="52%"}
**In R with `testthat`:**

```r
test_that("my_function is correct", {
  expect_equal(my_function(input), known_output)
  expect_error(my_function(bad_input))
})
```

**Test-driven development (TDD):**

1. **Red** — write a failing test for the behaviour you want
2. **Green** — write the code that makes it pass
3. **Refactor** — clean up; tests still pass

The agent *iterates against your tests* —
without tests, it has no objective criterion.
:::
::: {.column width="48%"}
**For statisticians: what to test**

- **Correctness**: numbers you can work out by hand (simple DGMs, closed-form results)
- **Boundary conditions**: edge cases, empty inputs, length-one vectors
- **Statistical properties**: symmetry, monotonicity, bounds on estimators
- **Reproducibility**: same seed → same result

::: {.callout-tip icon=false}
Tests are not bureaucracy —
they are the contract between
you and the code.
:::
:::
:::

::: notes
~2 min. The audience writes a lot of simulation code — this is where testthat is most valuable.
Emphasise that for statistical code you can often construct hand-checkable truth from closed-form expressions.
Reproducibility tests with set.seed() are trivially easy and catch a whole class of silent bugs.
:::

## Illustration (1/2) — Ask: write the function {.smaller}

**Prompt quality determines output quality.**

::: {.columns}
::: {.column width="55%"}
**A rigorous prompt:**

> *"Write an R function `coverage(truth, lower, upper)` that computes the empirical coverage probability of a set of confidence intervals. `truth` is a scalar, `lower` and `upper` are equal-length numeric vectors. Return the proportion of intervals that contain `truth`. Validate that `lower` and `upper` have equal length and that `lower <= upper` element-wise. Return a scalar between 0 and 1."*

**Generated output:**

```r
coverage <- function(truth, lower, upper) {
  stopifnot(length(lower) == length(upper))
  stopifnot(all(lower <= upper))
  mean(lower <= truth & truth <= upper)
}
```
:::
::: {.column width="45%"}
::: {.callout-important icon=false}
**Generated code is a hypothesis,
not a result.**

Read it · understand it · test it.

Would you submit a simulation result
without checking the DGM?
:::

**What made this prompt good?**

- Named the function and arguments
- Specified the return type
- Specified validation behaviour
- Specified the estimand (proportion)

*Vague prompt → vague output*
:::
:::

::: notes
~2 min. Walk through why the prompt is specific. Compare to a lazy prompt: "write a coverage function in R."
The coverage function is small enough that the audience can read and understand it in 30 seconds — that is the point. With a complex function you still need to understand it before trusting it.
:::

## Illustration (2/2) — Test it: validity and correctness {.smaller}

**Tests make the agent's loop self-correcting.**

::: {.columns}
::: {.column width="55%"}
```r
library(testthat)

test_that("coverage() is correct", {
  # Correctness — worked out by hand
  expect_equal(coverage(0, c(-1, -1), c(1,  1)), 1)   # all cover
  expect_equal(coverage(0, c( 1,  2), c(3,  4)), 0)   # none cover
  expect_equal(coverage(0, c(-1,  1), c(1,  2)), 0.5) # half cover

  # Boundary — closed endpoint counts
  expect_equal(coverage(1, 1, 2), 1)

  # Validity — bad input rejected
  expect_error(coverage(0, c(1, 2), 3))
})
```
:::
::: {.column width="45%"}
**The TDD loop with an agent:**

```
Write tests → agent writes function
→ tests FAIL (Red)
→ agent revises
→ tests PASS (Green)
→ you review diff and commit
```

::: {.callout-tip icon=false}
The test numbers are **yours** —
you worked them out by hand.
The agent cannot fake correctness
against truth you define.
:::
:::
:::

::: notes
~2 min. Walk through each test: the first three cover the logic, the boundary test checks edge behaviour, the error test checks validation.
The key point: the agent can iterate against these tests autonomously, but the truth values in the tests come from the statistician — that's the irreducible human contribution.
:::

# Worked example: simulation {background-color="#002546"}

## Double robustness under misspecification {.smaller}

**The question:**

> *How do the bias and variance of a doubly-robust GLM estimator compare to non-DR estimators when **both** the outcome model and the exposure model are misspecified?*

::: {.columns}
::: {.column width="55%"}
**Why this is the stress test**

A DR estimator is consistent if *either* nuisance model is correct.
Breaking both models removes the safety net entirely.

**Estimand:** marginal causal contrast

$$\tau = \mathbb{E}[Y(1)] - \mathbb{E}[Y(0)]$$

Binary outcome, binary exposure, confounders Z.

**Tool:** `stdReg2::standardize_glm_dr()`
*(Sachs, Sjölander, Gabriel, Ohlendorff & Brand)*
:::
::: {.column width="45%"}
**Competing estimators**

| Estimator | Outcome model | PS model |
|---|---|---|
| DR standardisation | misspecified | misspecified |
| Non-DR standardisation | misspecified | — |
| IPW | — | misspecified |

**DGM:** confounder Z with nonlinear
`X|Z` and `Y|X,Z` — linear working models are wrong by construction.

**Performance:** bias · variance · RMSE
· CI coverage · Monte Carlo SEs
:::
:::

::: notes
~3 min. Introduce the scientific question first, then the simulation design. The audience knows causal inference — lean on their knowledge.
stdReg2 is a package from this group, which makes it a natural centrepiece. Mention the binary-exposure requirement.
"Dual misspecification" is the stress test: if we break both models, how bad is it? The simulation will quantify that.
:::

## Build it with ADEMP — and let AI accelerate {.smaller}

**ADEMP** (Morris, White & Crowther, *Stat Med* 2019): a principled framework for simulation design.

| Element | What **you** decide | What **AI** accelerates |
|---|---|---|
| **Aims** | Compare DR vs non-DR under dual misspecification | Drafts study skeleton and runner |
| **Data-generating** | Confounder Z; nonlinear `X\|Z` and `Y\|X,Z` so linear working models are wrong | Writes the simulator — you verify it hits the true estimand |
| **Estimand** | Marginal contrast $\mathbb{E}[Y(1)]-\mathbb{E}[Y(0)]$, known from DGM | Computes the true value numerically |
| **Methods** | DR `standardize_glm_dr` (both models misspecified) vs non-DR / IPW | Wires up the competing method calls |
| **Performance** | Bias, variance, RMSE, CI coverage + Monte Carlo SEs | Codes the metrics — you insist on the MCSEs |

**Toolkit:** `SimDesign` · `rsimsum` · `future`/`furrr` · `set.seed()`

::: notes
~3 min. Walk the table column by column. The left column is intellectual work — ADEMP forces you to be explicit about each decision before you write code. The right column is mechanical — fast to generate, slow to verify.
Emphasise Monte Carlo SEs: they are easy to add and the single most common missing element in simulation studies.
:::

## The skill is in the prompt {.smaller}

::: {.columns}
::: {.column width="50%"}
**Lazy prompt:**

> *"Write me a simulation study comparing doubly-robust and non-doubly-robust estimators."*

**What you get:**

- Generic DGM with no meaningful misspecification
- No estimand stated
- No performance measures beyond bias
- No Monte Carlo SEs
- Code you cannot easily verify

::: {.callout-warning icon=false}
Garbage in → convincing-looking garbage out.
The model will produce *something* —
it will look complete and plausible.
:::
:::
::: {.column width="50%"}
**Rigorous ADEMP-structured prompt:**

> *"Design a simulation study using ADEMP. Aim: compare `stdReg2::standardize_glm_dr()` against non-DR standardisation and IPW for estimating $\mathbb{E}[Y(1)]-\mathbb{E}[Y(0)]$. DGM: binary Y, binary X, continuous confounder Z; use nonlinear relationships so linear working models are misspecified for both outcome and PS. Estimand: marginal risk difference. Methods: the three above, all using misspecified models. Performance: bias, empirical variance, RMSE, nominal-95% CI coverage — all with Monte Carlo SEs. Use `SimDesign` with `set.seed(42)`."*

**What you get:**

- Structured, reviewable code
- Correct estimand and DGM
- MCSEs included automatically
- Something you can actually critique
:::
:::

::: notes
~2 min. This is the most practically impactful slide in the deck. The contrast between the two prompts is stark.
Walk through what makes the good prompt good: it maps directly to ADEMP, it names the tool, it specifies misspecification explicitly, it demands MCSEs.
The model cannot improve the quality of your scientific decisions — it can only implement them faithfully.
:::

# Stages 3–4 + Govern {background-color="#002546"}

## Validate & analyse {.smaller}

**AI speeds the mechanics. Judgement stays human.**

::: {.columns}
::: {.column width="55%"}
**Where AI helps**

- Interpreting R/Python output: *"what does this warning mean?"*
- Running and explaining model diagnostics
- Explaining a method back to you in plain language
  *(a surprisingly useful cross-check)*
- Suggesting additional sensitivity analyses
- Formatting simulation results with `rsimsum`

:::
::: {.column width="45%"}
::: {.callout-important icon=false}
**Reproduce the headline numbers**

Before accepting AI-assisted output,
compute the key numbers yourself —
in a fresh script, from the raw results.

If the numbers disagree, investigate.
:::

::: {.callout-warning icon=false}
**Quiet deskilling**

If the AI always explains the output,
you stop reading it directly.
Stay engaged — use AI to *check*
your interpretation, not replace it.
:::
:::
:::

::: notes
~2 min. The "explain the method back" trick is underused: ask the model to explain what the function you just wrote actually does — it often surfaces misunderstandings.
Reproduce the headline numbers is the operational equivalent of the ground rule. One independent calculation is a minimum bar.
:::

## Translate & communicate — to the clinic {.smaller}

::: {.columns}
::: {.column width="55%"}
**Where AI helps**

- Draft and refine methods sections
- Tailor language for a clinical vs. statistical audience
- Respond to reviewer comments *(draft, then edit — never send raw)*
- Structure grant text against a funder's criteria
- Check a clinical prediction model report against guidelines

**AI reporting guideline checklist:**

| Guideline | Scope |
|---|---|
| TRIPOD+AI | prediction model development & validation |
| PROBAST-AI | prediction model risk-of-bias assessment |
| DECIDE-AI | decision support / clinical evaluation |
| CONSORT-AI | randomised trials with AI interventions |
:::
::: {.column width="45%"}
::: {.callout-important icon=false}
**The methodologist guarantees correctness**

AI can improve clarity and structure.
It cannot guarantee the methods
are correct for your research question.

*That is your job.*
:::

::: {.callout-tip icon=false}
**Practical tip**

Ask the model to evaluate your
methods section against TRIPOD+AI
checklist items — it will surface
gaps you missed.
:::
:::
:::

::: notes
~2 min. The clinical audience point is important for SMARTbiomed specifically — the pipeline ends at the clinic, and language mismatches cause real translation failures.
Reviewer responses: the model is genuinely useful for drafting, but you must edit carefully. Reviewers can often detect boilerplate.
The reporting guidelines checklist is a high-value, low-risk use case — the AI can do the tedious line-by-line check.
:::

## Govern: use it responsibly {.smaller}

::: {.columns}
::: {.column width="50%"}
**Privacy**

- **Never paste patient-level data** into public chat interfaces
- Use approved / on-premise deployments for sensitive data (check with your IT/data protection officer)
- GDPR applies to identifiable data in prompts — when in doubt, ask before pasting

**Reproducibility**

- Log the **prompts** you used for any analysis — treat them like code
- Record **tool versions** (model ID, package version)
- Your methods section should be reproducible from prompts + code + seed

:::
::: {.column width="50%"}
**Verification**

- Every **citation** — verify against the real source
- Every **number** — reproduce independently
- Every **derivation** — check the algebra yourself
- Every **API call** — check the documentation

**Disclosure**

- Most journals now require disclosure of AI use in writing or analysis
- ICMJE: authors are responsible for all AI-assisted content
- Check your target journal's policy *before* submission

::: {.callout-tip icon=false}
A one-paragraph "AI use statement" at
submission is becoming standard practice.
Write it as you go, not at the end.
:::
:::
:::

::: notes
~2 min. Privacy is the highest-stakes failure mode — one incident can damage trust in AI across the whole team.
Reproducibility logging is where most researchers fall short. Prompts are methods and should be treated as such.
The ICMJE point is practically important: journals are moving fast here and policies vary. Check early.
:::

# Closing {background-color="#002546"}

## Three things to try this week

::: {.columns}
::: {.column width="33%"}
**1. Install `gander`**

Try it in RStudio on a real dataset.
Ask it to write one function
using your actual column names.

*Lowest friction entry point
into in-IDE AI assistance.*
:::
::: {.column width="33%"}
**2. Draft a simulation with ADEMP**

Write out the five ADEMP elements
in plain text first.
Then paste them into a
structured prompt.
Verify the DGM hits the true estimand.

*The structure is the value —
not just for AI.*
:::
::: {.column width="33%"}
**3. Set one team ground rule**

Privacy: *"No patient-level data
in public chat interfaces."*

Disclosure: *"Log the prompts
used in any analysis."*

*A shared norm beats
individual policy every time.*
:::
:::

::: notes
~1 min. These are deliberately small and actionable. The goal is to lower the barrier to first use while anchoring good habits.
:::

## Over to you — 15 minutes {.smaller}

**Topic dot-vote** *(pick your top 2 — [Slido / stickers])*

::: {.columns}
::: {.column width="50%"}
- Setting up git for an existing project
- Writing your first unit tests in `testthat`
- Evaluating a DR simulation with `rsimsum`
:::
::: {.column width="50%"}
- Privacy and governance in your team's workflow
- Reporting guidelines for clinical AI
- Live demo: `gander` / Claude Code in action
:::
:::

---

**Think–pair–share prompt:**

> *Name one task you would hand to AI tomorrow —
> and one task you would never hand to AI.*

Share with a neighbour (2 min), then we'll hear a few pairs.

---
## Thank you — let's discuss {.smaller}

*Open floor — questions, pushback, and reality checks welcome.*

::: notes
Facilitator: run the dot-vote first to decide which topic gets the deep dive. The think-pair-share can run in parallel or after.
Leave room for sceptical questions — they are the most useful ones.
:::


