---
title: "Open Source Software"
author: "Michael C Sachs"
date: "Updated `r format(Sys.Date(), '%Y-%m-%d')`"
output: 
  html_document: 
    css: assets/style.css
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(gh)
library(jsonlite)
#library(icon)
library(htmltools)
```

<style type="text/css">

#title {
    color: #e9e9e9;
	padding: 0.8em;
    font-family: Open sans, sans-serif;
    background-color: #dc4d38;
	font-size: 32px;
	text-align: center;

	-moz-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	-webkit-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	margin-bottom: 20px;
	padding: 0.4em;
	width: 80%;

	display: table;
	position: relative;
	margin: auto;
	margin-top: 20px;
	margin-bottom: 20px;
}


.email {
   	font-size: 26px;
}

.stats {
	color: white;
	padding: 0.4em;
	font-family: Open sans, sans-serif;
	background: #39b1a4;

	font-size: 26px;
	text-align: center;

	-moz-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	-webkit-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	margin-bottom: 20px;
	padding: 0.4em;
	width: 65%;

	display: table;
	position: relative;
	margin: auto;
	margin-top: 20px;
	margin-bottom: 20px;
}

ol {
display: block;
list-style-type: none;
-webkit-margin-before: 0em;
-webkit-margin-after: 0em;
-webkit-margin-start: 0px;
-webkit-margin-end: 0px;
-webkit-padding-start: 0px;
}

.repolist {

	max-width: 1096px;
	margin: auto;

}

.repo {
	background: white;
	-moz-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	-webkit-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	margin-bottom: 20px;
	padding: 1em;
	width: inherit;

	overflow: hidden;

	position: relative;
	margin-left: 10px;
	margin-right: 10px;
}

.repotitle {
	font-size: 26px;
	text-align: center;

	background: white;
	-moz-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	-webkit-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	margin-bottom: 20px;
	padding: 0.4em;
	width: 90%;

	display: table;
	position: relative;
	margin: auto;
	margin-top: 20px;
	margin-bottom: 20px;

}

.reponame {
	font-size: 22px;
	text-align: center;
	color: #006FA3;
}

</style>

```{r repos, echo = FALSE}
repos <- gh("/users/sachsmc/repos", .token = readLines(".mtok")[1], .limit = 65)

test <- fromJSON(toJSON(repos))

myreps <- subset(test, fork == FALSE & private == FALSE)[, c("name", "description", "html_url", 
                                                   "stargazers_count", "watchers_count", "language", "has_pages", 
                                                   "updated_at", "pushed_at")]

makecard <- function(row, set) {
    
    tags$li(class = "repo", 
            div(
                a(class = "reponame", href = set$html_url[row], set$name[row]), 
                h3(set$language[row]), 
                p(set$description[row]), 
                p(class = "stargazers", paste(set$stargazers_count[row], "stars")), 
                a(href = ifelse(set$has_pages[row] == TRUE, paste0("https://sachsmc.github.io/", set$name[row]), 
                                set$html_url[row]), "Project homepage")
            ))
    
}
```


# Recently Updated
```{r}

newr <- myreps[order(as.Date(unlist(myreps$pushed_at)), decreasing = TRUE), ][1:10, ]

listrp <- lapply(1:nrow(newr), makecard, newr)
tags$ol(
    listrp
)

```


# Popular
```{r}

popl <- myreps[order(unlist(myreps$stargazers_count), decreasing = TRUE), ][1:10, ]

listrp <- lapply(1:nrow(popl), makecard, popl)
tags$ol(
    listrp
)

```


# R packages
```{r}

popl <- subset(myreps, language == "R")
popl <- popl[order(as.Date(unlist(popl$updated_at)), decreasing = TRUE), ]

listrp <- lapply(1:nrow(popl), makecard, popl)
tags$ol(
    listrp
)

```
