---
title: "Recent Presentations"
author: "Michael C Sachs"
date: "Updated `r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Recent presentations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


<style type="text/css">

#title {
    color: #e9e9e9;
	padding: 0.8em;
    font-family: Open sans, sans-serif;
    background-color: #dc4d38;
	font-size: 32px;
	text-align: center;

	-moz-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	-webkit-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	margin-bottom: 20px;
	padding: 0.4em;
	width: 80%;

	display: table;
	position: relative;
	margin: auto;
	margin-top: 20px;
	margin-bottom: 20px;
}


.email {
   	font-size: 26px;
}

.stats {
	color: white;
	padding: 0.4em;
	font-family: Open sans, sans-serif;
	background: #39b1a4;

	font-size: 26px;
	text-align: center;

	-moz-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	-webkit-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	margin-bottom: 20px;
	padding: 0.4em;
	width: 65%;

	display: table;
	position: relative;
	margin: auto;
	margin-top: 20px;
	margin-bottom: 20px;
}

ol {
display: block;
list-style-type: none;
-webkit-margin-before: 0em;
-webkit-margin-after: 0em;
-webkit-margin-start: 0px;
-webkit-margin-end: 0px;
-webkit-padding-start: 0px;
}

.repolist {

	max-width: 1096px;
	margin: auto;

}

.repo {
	background: white;
	-moz-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	-webkit-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	margin-bottom: 20px;
	padding: 1em;
	width: inherit;

	overflow: hidden;

	position: relative;
	margin-left: 10px;
	margin-right: 10px;
}

.repotitle {
	font-size: 26px;
	text-align: center;

	background: white;
	-moz-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	-webkit-box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	box-shadow: 0px 0px 5px 1px rgba(0, 0, 0, 0.1);
	margin-bottom: 20px;
	padding: 0.4em;
	width: 90%;

	display: table;
	position: relative;
	margin: auto;
	margin-top: 20px;
	margin-bottom: 20px;

}

.reponame {
	font-size: 22px;
	text-align: center;
	color: #006FA3;
}

</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rmarkdown)
library(htmltools)

tkfold <- list.dirs("../talks", recursive = FALSE, full.names = FALSE)

process_talk <- function(dir) {
    
    linkrmd <- file.path("../articles/talks", dir, "index.Rmd")
    titles <- yaml_front_matter(linkrmd)
    
    tags$li(class = "repo", 
            div(
                h3(a(class = "reponame", href = gsub(".Rmd", ".html", linkrmd, fixed = TRUE), titles$title)), 
                p(titles$date), 
                p(titles$author)
            ))
    
}

```


```{r}
listrp <- lapply(tkfold, process_talk)
tags$ol(
    listrp
)
```
